\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-08-20 onebit_full.tex
added a note on the extension of Thm. 2 to other loss functions

\subsection{Multiterminal Source Coding \label{sec:ceo}}
The CEO setting considers the estimation of a sequence $\theta_1,\theta_2\ldots$, where a noisy version of each $\theta_i$ is available at $n$ terminals. At each terminal $i$, an encoder observes the $k$ noisy samples
\[
X_{i,j} = \theta_j + Z_{i,j},\qquad j=1,\ldots,k, \qquad i = 1,\ldots,n,
\]
and transmits $r_i k$ bits to a central estimator  \cite{berger1996ceo}. The central estimator produces an estimates ${\theta}_1,\ldots,{\theta}_k$ with the goal of minimizing the quadratic risk:
\[
R_{\CEO} = \frac{1}{k} \sum_{j=1}^k \mathbb E \left[\left(\theta_j - {\theta}_j \right)^2 \right]. 
\]
Note that any distributed encoding scheme using one-bit per sample can be replicated $k$ times and thus leads to a legitimate encoding and estimation scheme for the CEO problem with $r_1=\ldots=r_n = 1$. It follows that, assuming that $\theta$ is drawn once from the prior $\pi(d\theta)$, our mean estimation problem from one-bit samples under distributed encoding corresponds to the CEO setting with $k=1$ realization of $\theta$ observed under noise at $n$ different locations, and communicated at each location using an encoder sending a single bit. 
%
Consequently, a lower bound on the MSE in estimating $\theta$ in the distributed encoding setting is given by the minimal MSE in the CEO setting as $k \to \infty$. Note that the difference between the CEO setting and ours lays in the privilege of each of the encoders to describe $k$ realizations of $\theta$ using $k$ bits with MSE averaged over these realizations, rather than a single realization using a single bit in ours. 
 \par
When the prior on $\theta$ and the noise corrupting it at each location are Gaussian, the optimal encoding scheme and its asymptotic risk as $k$ goes to infinity were fully characterized in \cite{prabhakaran2004rate}. Furthermore, the work of \cite{chen2004upper} provided an expression for quadratic risk attained in the CEO setting under Gaussian priors. Adapting to our setting, this expression provides the following proposition:
\begin{prop} \label{prop:ceo_lower_bound}
Assume that $\Theta = \mathbb R$ and $\pi(\theta) = \Ncal(0,\sigma_\theta^2)$ where $\sigma_\theta^2 \in \mathbb R$ is arbitrary. Then any estimator ${\theta}_n$ of $\theta$ in the distributed setting satisfies
\begin{equation} \label{eq:ceo_bound}
 n\mathbb E \left( \theta - \theta_n \right)^2 \geq \frac{4}{3} \sigma^2 + O(n^{-1}),
\end{equation}
where the expectation is with respect to $\theta$ and $X^n$.
\end{prop}

@@

Theorem~\ref{thm:adpative_lower_bound} implies that any estimator ${\theta}_n$ from any adaptive encoding scheme satisfies
\[
n\mathbb E\left[ (\theta-\theta_n)^2 \right] \leq  \frac{1}{4f^2(0)}+O(n^{-1}),
\]
and 
\[
\ARE({\theta}_n) \leq 4f^2(0)\sigma^2 = \eta(0)\sigma^2.
\]

Next, we present an adaptive encoding and estimation scheme that attains the maximal ARE of $\eta(0)\sigma^2$.

>>>>

\subsection{Multiterminal Source Coding \label{sec:ceo}}
The CEO setting considers the estimation of a sequence $\theta_1,\theta_2\ldots$, where a noisy version of each $\theta_i$ is available at $n$ terminals. At each terminal $i$, an encoder observes the $k$ noisy samples
\[
X_{i,j} = \theta_j + Z_{i,j},\qquad j=1,\ldots,k, \qquad i = 1,\ldots,n,
\]
and transmits $r_i k$ bits to a central estimator  \cite{berger1996ceo}. The central estimator produces an estimates ${\theta}_1,\ldots,{\theta}_k$ with the goal of minimizing the quadratic risk:
\[
R_{\CEO} = \frac{1}{k} \sum_{j=1}^k \mathbb E \left[\left(\theta_j - {\theta}_j \right)^2 \right]. 
\]
Note that any distributed encoding scheme using one-bit per sample can be replicated $k$ times and thus leads to a legitimate encoding and estimation scheme for the CEO problem with $r_1=\ldots=r_n = 1$. It follows that, assuming that $\theta$ is drawn once from the prior $\pi(d\theta)$, our mean estimation problem from one-bit samples under distributed encoding corresponds to the CEO setting with $k=1$ realization of $\theta$ observed under noise at $n$ different locations, and communicated at each location using an encoder sending a single bit. 
%
Consequently, a lower bound on the MSE in estimating $\theta$ in the distributed encoding setting is given by the minimal MSE in the CEO setting as $k \to \infty$. Note that the difference between the CEO setting and ours lays in the privilege of each of the encoders to describe $k$ realizations of $\theta$ using $k$ bits with MSE averaged over these realizations, rather than a single realization using a single bit in ours. 
 \par
When the prior on $\theta$ and the noise corrupting it at each location are Gaussian, the optimal encoding scheme and its asymptotic risk as $k$ goes to infinity were fully characterized in \cite{prabhakaran2004rate}. Furthermore, the work of \cite{chen2004upper} provided an expression for quadratic risk attained in the CEO setting under Gaussian priors. Adapting to our setting, this expression provides the following proposition:
\begin{prop} \label{prop:ceo_lower_bound}
Assume that $\Theta = \mathbb R$ and $\pi(\theta) = \Ncal(0,\sigma_\theta^2)$ where $\sigma_\theta^2 \in \mathbb R$ is arbitrary. Then any estimator ${\theta}_n$ of $\theta$ in the distributed setting satisfies
\begin{equation} \label{eq:ceo_bound}
 n \ex{\left( \theta - \theta_n \right)^2} \geq \frac{4}{3} \sigma^2 + O(n^{-1}),
\end{equation}
where the expectation is with respect to $\theta$ and $X^n$.
\end{prop}

@@

Theorem~\ref{thm:adpative_lower_bound} implies that any estimator ${\theta}_n$ from any adaptive encoding scheme satisfies
\[
n\mathbb E\left[ (\theta-\theta_n)^2 \right] \geq  \frac{1}{4f^2(0)}+O(n^{-1}),
\]
and 
\[
\ARE({\theta}_n) \leq 4f^2(0)\sigma^2 = \eta(0)\sigma^2.
\]
It is possible to extend Theorem~\ref{thm:adpative_lower_bound} to any other loss functions using a more subtle version of the Van Tress inequality provided in \cite{efroimovich1980information}. See also \cite{DBLP:journals/corr/abs-1902-08582}. \par
%
Next, we provide an adaptive encoding and estimation scheme that attains the maximal ARE of $\eta(0)\sigma^2$.
